Variable Selection

### 2015-06-24 var_select.rmd
revised core df to exclude client and contact codes

# library('knitr', lib = 'C:/Progra~1/R/R-3.1.2/library')
# library('ggplot2', lib = 'C:/Progra~1/R/R-3.1.2/library')
# library("dplyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
# library("plyr",lib = 'C:/Progra~1/R/R-3.1.2/library')
# library('magrittr',lib='C:/Progra~1/R/R-3.1.3/library')
# library('reshape2',lib='C:/Progra~1/R/R-3.1.3/library')
# library("rpart",lib = 'C:/Program Files/R/R-3.2.0/library')
# library('car', lib = 'C:/Progra~1/R/R-3.2.0/library')
# library('e1071', lib = 'C:/Progra~1/R/R-3.2.0/library')
# library('corrgram', lib = 'C:/Progra~1/R/R-3.2.0/library')
# library('party', lib = 'C:/Progra~1/R/R-3.2.0/library')

library('knitr')
library('ggplot2')
library("dplyr")
library("plyr")
library('magrittr')
library('reshape2')
library("rpart")
library('car')
library('e1071')
library('corrgram')
library('party')

# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
```

@@

@@

```{r core_creation}
#impute NA values in some variables to be in core, call this all7b
all7b<- all7a
all7b[is.na(all7b$State),]$State <- 'QLD'
#RSL Care client - now have details
all7b[all7b$code.client %in% 'C2835',]$no.employees <- 1001
all7b[all7b$code.client %in% 'C2835',]$Biz.type <- 'NFP'
all7b[all7b$code.client %in% 'C2835',]$Business <- 'developer/real estate'
all7b[all7b$code.client %in% 'C2758',]$Biz.size <- 'national'
all7b[all7b$code.client %in% 'C2758',]$no.employees <- 2
all7b[all7b$code.client %in% 'C2758',]$Biz.size <- 'local'
all7b[all7b$code.client %in% 'C2758',]$Business <- 'business'
all7b[all7b$code.client %in% 'C2326',]$Business <- 'health'
all7b[all7b$code.client %in% 'C2338',]$Biz.size <- 'local'
all7b[all7b$code.client %in% 'C2338',]$Business <- 'developer/real estate'
all7b[all7b$code.client %in% 'C2338',]$Biz.type <- 'private'
all7b[all7b$code.client %in% 'C2882',]$Biz.size <- 'local'
all7b[all7b$code.client %in% 'C2882',]$Business <- 'builder'
all7b[all7b$code.client %in% 'C2882',]$Biz.type <- 'private'
all7b[all7b$code.client %in% 'C2882',]$no.employees <- 3
#all remaining NA Business will be 'unknown'
all7b$Business<- as.character(all7b$Business)
all7b[is.na(all7b$Business),]$Business <- 'unknown'
all7b$Business<- as.factor(all7b$Business)

#impute Biz.size to be 'local' by default - this is most popular category
all7b[is.na(all7b$Biz.size),]$Biz.size <- 'local'
#impute Biz.type to be 'private' by default - this is most popular category
#client which is 'department of national parks' shall be 'gov'
all7b[is.na(all7b$Biz.type),]$Biz.type <- 'private'
all7b[all7b$code.client %in% 'C2503',]$Biz.type <- 'gov'
#delete charge.ph and charge.pc, Start.Date and End.Date
all7b<- all7b %>% select(-charge.ph, -charge.pc, -Start.Date, -End.Date)
#impute NA no.users as 2 = median and mode
all7b[is.na(all7b$no.users),]$no.users <- 2
#impute NA num.disc as 1 - overwhelmingly the most common
all7b[is.na(all7b$Num.disc),]$Num.disc <- 1
#num.inv and mean.inv both have a single NA value
# this consisted of single large -ve invoice. I think we should delete this case 
all7b<- all7b %>% filter(!is.na(num.inv))
# write.csv(all7b,'all7b.csv')

@@

var.imp<- cforest(return.pdol~.,
        data= all7b
        )

@@

>>>>

library('knitr', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('ggplot2', lib = 'C:/Progra~1/R/R-3.2.1/library')
library("dplyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library("plyr",lib = 'C:/Progra~1/R/R-3.2.1/library')
library('magrittr',lib='C:/Progra~1/R/R-3.2.1/library')
library('reshape2',lib='C:/Progra~1/R/R-3.2.1/library')
library("rpart",lib = 'C:/Program Files/R/R-3.2.1/library')
library('car', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('e1071', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('corrgram', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('party', lib = 'C:/Progra~1/R/R-3.2.1/library')
library('randomForest', lib = 'C:/Progra~1/R/R-3.2.1/library')

# library('knitr')
# library('ggplot2')
# library("dplyr")
# library("plyr")
# library('magrittr')
# library('reshape2')
# library("rpart")
# library('car')
# library('e1071')
# library('corrgram')
# library('party')

# setwd("~/OneDrive/shared files/Bligh Tanner/masters/data")
# setwd("C:/Users/n9232371/Documents/Consultbusiness/data")
opts_knit$set(root.dir= "~/OneDrive/shared files/Bligh Tanner/masters/data")
```

@@

# write.csv(all7a,'all7a.csv')
all7a<- read.csv('all7a.csv')
```

@@

```{r core_creation, include=FALSE}
#impute NA values in some variables to be in core, call this all7b
all7b<- all7a
all7b[is.na(all7b$State),]$State <- 'QLD'
#RSL Care client - now have details
all7b[all7b$code.client %in% 'C2835',]$no.employees <- 1001
all7b[all7b$code.client %in% 'C2835',]$Biz.type <- 'NFP'
all7b[all7b$code.client %in% 'C2835',]$Business <- 'developer/real estate'
all7b[all7b$code.client %in% 'C2758',]$Biz.size <- 'national'
all7b[all7b$code.client %in% 'C2758',]$no.employees <- 2
all7b[all7b$code.client %in% 'C2758',]$Biz.size <- 'local'
all7b[all7b$code.client %in% 'C2758',]$Business <- 'business'
all7b[all7b$code.client %in% 'C2326',]$Business <- 'health'
all7b[all7b$code.client %in% 'C2338',]$Biz.size <- 'local'
all7b[all7b$code.client %in% 'C2338',]$Business <- 'developer/real estate'
all7b[all7b$code.client %in% 'C2338',]$Biz.type <- 'private'
all7b[all7b$code.client %in% 'C2882',]$Biz.size <- 'local'
all7b[all7b$code.client %in% 'C2882',]$Business <- 'builder'
all7b[all7b$code.client %in% 'C2882',]$Biz.type <- 'private'
all7b[all7b$code.client %in% 'C2882',]$no.employees <- 3
#all remaining NA Business will be 'unknown'
all7b$Business<- as.character(all7b$Business)
all7b[is.na(all7b$Business),]$Business <- 'unknown'
all7b$Business<- as.factor(all7b$Business)

#impute Biz.size to be 'local' by default - this is most popular category
all7b[is.na(all7b$Biz.size),]$Biz.size <- 'local'
#impute Biz.type to be 'private' by default - this is most popular category
#client which is 'department of national parks' shall be 'gov'
all7b[is.na(all7b$Biz.type),]$Biz.type <- 'private'
all7b[all7b$code.client %in% 'C2503',]$Biz.type <- 'gov'
#delete charge.ph and charge.pc, Start.Date and End.Date
all7b<- all7b %>% select(-charge.ph, -charge.pc, -Start.Date, -End.Date)
#impute NA no.users as 2 = median and mode
all7b[is.na(all7b$no.users),]$no.users <- 2
#impute NA num.disc as 1 - overwhelmingly the most common
all7b[is.na(all7b$Num.disc),]$Num.disc <- 1
#num.inv and mean.inv both have a single NA value
# this consisted of single large -ve invoice. I think we should delete this case 
all7b<- all7b %>% filter(!is.na(num.inv))
# 7 NA cases with significant dis.sc.mlsto - ie done by contractor 
pc<- c('pc.contracttech','pc.director', 'pc.midtech', 'pc.midpro', 'pc.gradpro', 'pc.seniortech', 'pc.seniorpro', 'pc.pro', 'pc.tech')

for (i in 1:length(pc)){
        all7b[is.na(all7b[,pc[i]]),][,pc[i]] <-0
}

#change unknown Num.disc to be 1
all7b[is.na(all7b$Num.disc),]$Num.disc <- 1
# write.csv(all7b,'all7b.csv')

Start running cforests
Very slow, start with 100 rows only, using only core variables

```{r, echo=FALSE}
all7b<- read.csv('all7b.csv')

@@

mini.core2 <- all7b %>% select(code.client , Discipline ,  client.count , inv.mlsto , Business ,
        Biz.size , Biz.type , no.users , timespan, Num.disc, 
        return.pdol, code.contact,
        pc.director,
        pc.pro , code.director , ProjEng.Pos ,
        client.meaninv , client.neginv , client.numinv , client.totinv)

mini.core3 <- all7b %>%
        select(Discipline , inv.mlsto , Business,
        Biz.size , Biz.type , no.users , timespan, Num.disc, 
        return.pdol, pc.director,
        pc.pro , code.director , ProjEng.Pos,
        client.meaninv , client.neginv , client.numinv , client.totinv)

@@

Try to make all cforests random selection of 200 rows - this time post 2009

Try running a cforest over night with full dataset - don't include 
```

system.time(
all.cfor2<- cforest(return.pdol~.,
        data= mini.core3
        )
)

var.imp4<- varimp(all.cfor) %>% as.data.frame
colnames(var.imp4)[names(var.imp4) %in% '.']<-'var.imp'
var.imp4$var <- rownames(var.imp4)
#reorder levels for bar plot
var.imp4 <- within(var.imp4, 
                    var <- factor(var,levels=
                                            var.imp4[order(-abs(var.imp4$var.imp)),]$var))

q<- ggplot(data=var.imp4, aes(x=var, y=var.imp)) + theme(axis.text.x=element_text(angle=45,hjust=1))
q + geom_bar(stat='identity', position = 'dodge')

save(all.cfor,file= 'C:/Users/n9232371/Documents/github/consult/object/all_cfor.RData')
load('C:/Users/n9232371/Documents/github/consult/object/all_cfor.RData')
```

From analysing all jobs including 2009 and beyond, the most important variables are:
`levels(var.imp4$var)[1:10] %>% as.data.frame`

The least important are 
`levels(var.imp4$var)[11:19]`

Let's compare this to

* all data from 2002 to 2014 - currently running in another R window
* ANOVA - all data because fast
* random forest

## Anova:

```{r anova_varimp, echo=FALSE}
#use mini.core2- all years
system.time(
        aov.all<- aov(return.pdol~.,
                      data= mini.core3)
        )

summary(aov.all)
```

Therefore most significant variables are:

* no.users - number of people on the job - <<.1%
* code.director - <<.1%
* timespan - 4%
* Num.disc - 5.6%
* Discipline - 5.65%

worst: Business, ProjEng.Pos, client.numinv,

Contrasts to cforest results include

* code.client pval=1 - this is likely because code.contact and code.client are highly correlated
* inv.mlsto pval= .721 - surprising
* client.numinv pval = 0.923 - surprisingly different

# random forests

First optimise parameters mtry and ntree - exclude code.client and code.contact for now as they have way too many categories

```{r randomForest, echo=FALSE}
#random forest can have max 53 categories in a variable
#code.client has the most categories - 664

# find optimal mtry

#create train/test sets

train_ind <- sample(seq_len(2355), size = 1570)

train <- mini.core4[train_ind, ]
test <- mini.core4[-train_ind, ]
oob.err<-double(13)
test.err<-double(13)
for(mtry in 1:13){
        fit<-randomForest(return.pdol ~., data=train, mtry=mtry, ntree=400)
        oob.err[mtry]<-fit$mse[400]
        test.err[mtry]<- mean((predict(fit,test)-test$return.pdol)^2)
        cat(mtry," ")
}

qplot(1:mtry,test.err,geom=c("point","line"),color="pink")+
        geom_line(aes(1:mtry,oob.err,colour="blue"))+
        geom_point(aes(1:mtry,oob.err, colour="blue"))

#COMPLETELY FLAT - MAY AS WELL USE MTRY = 5????? NOW VARY NUMBER OF TREES

trees<- c(300,500,700,1000,1500)
oob.err<-double(5)
test.err<-double(5)
for(ntree in 1:5){
        fit<-randomForest(return.pdol ~., data=train, mtry=5, ntree=trees[ntree])
        oob.err[ntree]<-fit$mse[trees[ntree]]
        test.err[ntree]<- mean((predict(fit,test)-test$return.pdol)^2)
        cat(ntree," ")
}

qplot(1:ntree,test.err,geom=c("point","line"),color="pink")+
        geom_line(aes(1:ntree,oob.err,colour="blue"))+
        geom_point(aes(1:ntree,oob.err, colour="blue"))

#again, doesn't matter!! line is dead flat. i'm not having much confidence in the predictiveness of this model. 
```

Extract important variables from randomForest loop

423 of the 664 client companies and 485 of the 670 client contacts appear ONLY once in the entire data set.
While they may be great predictors what do we do in the majority of cases where the contact is new!?

```{r, echo=FALSE}
set.seed(47)

#try running without code.client or code.contact
multi.forest1<-function(data=mini.core3, nforest=1:10){
        sample=double(50)
        imp = NULL
        for (i in 1:length(nforest)){

                #run tree
                assign(paste("Forest",i,sep=""),randomForest(return.pdol~.,
                                                             data=data,
                                                             ntree=400,mtry=5, importance=TRUE),
                       envir=.GlobalEnv)

                #put together importance rank for each variable

                rand.imp= importance(paste("Forest",i,sep="") %>% get, type=1, scale=FALSE) %>% as.data.frame
                colnames(rand.imp)[names(rand.imp) %in% '%IncMSE']= 'IncMSE'
                rand.imp$var= rownames(rand.imp)
                rand.imp= rand.imp %>% arrange(-abs(IncMSE))
                rand.imp$rank= 1:nrow(rand.imp)

                #rbind all variables and ranks on top of one another for each Forest

                imp = rbind(imp, rand.imp %>% select(var,rank))

                cat(i," ")

        #output average rank of each variable
        rank = ddply(imp, .(var), summarise, m.rank =mean(rank)) %>% arrange(m.rank)
        return(rank)

rand.rank<- multi.forest1(data=mini.core2 %>% select(-code.contact, -code.client), nforest=1:10)

kable(rand.rank)

See how variable selection goes with random forest first, then probably just run cforest again without code.client and contact.code
